### 2023-03-09 resilience.tex
Changed name of Uniform resilience to remove 'Uniform'

\section{Uniform resilience for WSTS}

@@

\problemx{resilience problem (URP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{*} Safe$ ?\newline}
%
%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}

\problemx{$k$-resilience problem (UkRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}

We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
%
\problemx{bounded resilience problem (UBRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is uniformely $k$-resilient ?\newline}

@@

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
%		
URP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

@@

\begin{proposition}\label{general}
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
\end{proposition}

@@

\begin{theorem}\label{up-down}
The three uniform resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
%	and bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is uniformely (bad,Safe)-resilient. Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm. 
%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$. similar reasonning for the other case.
hence the two uniform resilience problems are decidable.

Let us examine the uniform $k$-resilience problem (UkRP). we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
\end{proof}

@@

\begin{proposition}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
The uniform resilience problem (URP) and the uniform bounded resilience problem (UBRP) are decidable.
The uniform $k$-resilience problem (UkRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
\end{proposition}

\begin{proof}
Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
The uniform resilience problem (URP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$. This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.

@@

To decide the (uniform) bounded resilience problem (UBRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is uniformely $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the uniform resilience problem (UBRP) and the uniform bounded resilience problem (UBRP) are decidable.

To decide the UkRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is uniformely $k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. S is uniformely $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then UkRP is decidable.
\end{proof}

>>>>

\section{Resilience for WSTS}

@@

\problemx{resilience problem (RP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{*} Safe$ ?\newline}
%
%\alain{il faudrait ne pas répéter 3 fois les mêmes imputs pour les 3 pbs: énoncer les 3 uniformes pbs d'un coup avec une fois l'input puis les 3 pbs pour un état s donné sans répéter non plus 3 fois les mêmes inputs}

\problemx{$k$-resilience problem (kRP)}
{A transition system $\mathscr{S}=(S,\rightarrow), k \in \mathbb{N}$ and two sets $Safe, Bad \subseteq S$.}
{$Bad \longrightarrow^{\leq k} Safe$ ?\newline}

We add a third problem that decides whether there exists an $k$ such  that the system is $k$-resilient.
%
\problemx{bounded resilience problem (BRP)}
{A transition system $\mathscr{S}=(S,\rightarrow)$ and two sets $Safe, Bad \subseteq S$.}
%{$\exists k \geq 0 ~ \forall s' \in D ~ s \rightarrow^* s' \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
{$\exists k \geq 0$ such that $\mathscr{S}$ is %uniformely
 $k$-resilient ?\newline}

@@

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of Safe. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that Bad is always downward closed and Safe is always upward closed.
%		
RP is decidable for lossy counter machines (LCM) with Safe and Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

@@

\begin{proposition}\label{general}
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient iff $bad \subseteq \pred^*(Safe)$.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient.\\
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-$k$-resilient iff $bad \subseteq \pred^k(Safe)$.
\end{proposition}

@@

\begin{theorem}\label{up-down}
The three %uniform 
resilience problems are decidable for WSTS with effective predbasis and $Safe=\uparrow Safe$.
%	and bad is upward closed or downward closed.
\end{theorem}

\begin{proof}
Let us first solve if $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(bad,Safe)-resilient. Since $Safe=\uparrow Safe$, and WSTS, $\pred^*(Safe)=\uparrow \pred^*(Safe)$, so $\pred^*(Safe)$ admits a finite basis $B_{\pred^*(Safe)}$. since WSTS  with effective predbasis, we may compute a finite basis of $\pred^*(Safe)$ with the backward coverability algorithm. 
%Then if bad is upward closed or downward closed, we may decide the inclusion $Bad \subseteq \uparrow B_{\pred^*(Safe)}$ which remais to compare, in both cases, two finite sets. 
if bad  is upward closed, $bad = \uparrow B_{bad}$ and $ \uparrow B_{bad} \subseteq \uparrow B_{\pred^*(Safe)}$ iff for every $b \in B_{bad}$, there is a $s \in B_{\pred^*(Safe)}$ such that $s \leq b$. similar reasonning for the other case.
hence the two %uniform
 resilience problems are decidable.

Let us examine the %uniform 
$k$-resilience problem (kRP). we may compute the least $n$ such that $bad \subseteq \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely 
$k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
\end{proof}

@@

\begin{proposition}\label{down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and two finite sets: $I_{BAD}$ and $Min_{Safe}$.
The %uniform 
resilience problem (RP) and the %uniform 
bounded resilience problem (BRP) are decidable.
The %uniform
 $k$-resilience problem (kRP) is decidable if moreover $\mathscr{S}$ has the predbasis hypothesis.
\end{proposition}

\begin{proof}
Let $I_{BAD}=\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $BAD$ and $B_{Safe}=\{b_1,b_2,...,b_m\}$ be the minimal basis of $Safe$.
The %uniform 
resilience problem (RP) is equivalent to an infinite number of instances of the coverability problem : for all $x \in BAD$ does there exist an $j$ such that $b_j$ is coverable from $x$. This infinite set of coverability questions can be reduced to $n$ instances of the coverability problem in the completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ of $\mathscr{S}=(S,\rightarrow, \leq)$.

@@

To decide the %(uniform)
 bounded resilience problem (BRP), we decide $n \times m$ coverability questions: is state $b_j$ coverable from ideal $J_i$ ? If all these $n \times m$ coverability questions are positive then we compute $K=max(k_a \mid a=1,..,m \times n)$ and we deduce that  $\mathscr{S}$ is %uniformely
  $K$-resilient. Since coverability is decidable for completion-post-effective $\omega^2$-WSTS, we have shown that both the %uniform 
  resilience problem (BRP) and the %uniform 
  bounded resilience problem (BRP) are decidable.

To decide the kRP, we may compare the input $k$ with $K$. If $k \geq K$, we conclude that is %uniformely 
$k$-resilient; if $k < K$ we cannot answer. if moreover $\mathscr{S}=(S,\rightarrow, \leq)$ has the predbasis hypothesis, we may compute the least $n$ such that $ \uparrow \pred^n(Safe)=  \uparrow \pred^*(Safe)$. $S$ is %uniformely
 $k$-resilient iff $n \leq k$ \alain{hum...mais il peut y avoir des chemins plus courts}and then kRP is decidable.
\end{proof}
